---
title: "hw02_gapminder.Rmd"
author: "Fariha Khan"
date: '2017-09-24'
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




#### Work with the gapminder data we explored in class


Load Gapminder and dplyr (via tidyverse)

```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gapminder))
```

##### Explore the gapminder object:

1. Is it a data.frame, a matrix, a vector, a list?
- *The gapminder dataset is a list object and a data frame*
```{r}
str(gapminder)
```

2. What’s its class?
- *The gapminder dataset is a data frame class (made of lists)*
```{r}
class(gapminder)
typeof(gapminder)
```

3. How many variables/columns?
- *There are 6 column*
```{r}
ncol(gapminder)
```

4. How many rows/observations?
- *There are 1704 rows*
```{r}
nrow(gapminder)
```

5. Can you get these facts about “extent” or “size” in more than one way? Can you imagine different functions being useful in different contexts?
- *The str() function displays a summary of the internal structure of the object*
- *The summary() function shows the name and a summary of each variable*
- *The dim() function gives the value of rows and columns*
- *The length() function gives the number of columns and the names() fucntion outputs the names of of each column*
```{r}
str(gapminder)
summary(gapminder)
dim(gapminder)
names(gapminder)
length(gapminder)
```

6. What data type is each variable?
- *The str() function summaries the structure of the object and gives the variable type and range. The variable names and types are:*

       - country  : factor
       - continent: factor
       - year     : integer
       - lifeExp  : numerical
       - pop      : integer
       - gdpPercap: numerical
       
```{r}
str(gapminder)
```


#### Explore individual variables

##### Pick at least one categorical variable and at least one quantitative variable to explore.
I chose to explore explore the overall population by continent here.

First I used tally() to show the different columns and the number of observations in each conintent.
      
```{r tally continents}
gapminder %>% 
      group_by(continent) %>% 
      tally()
```

I used the summarize() to get the range of the minimum, maximum and average population observed in a given continent.
```{r}
gapminder %>% 
      group_by(continent) %>% 
      summarise(min_population = min(pop),
                avg_population = mean(pop),
                max_population = max(pop))
```

```{r, echo=FALSE}
continent_range <- gapminder %>% 
      group_by(continent) %>% 
      summarise(min_population = min(pop),
                avg_population = mean(pop),
                max_population = max(pop))
```

```{r, echo=FALSE, fig.height=7, fig.width=7}
View(continent_range)
```


Here I added a second categorical variable and looked to find which country had the lowest and highest population per given continent. 
```{r}
gapminder %>% 
      group_by(continent) %>%
      select(continent, country, pop) %>%
      top_n(1, wt = (pop)) %>% 
      arrange(continent)

gapminder %>% 
      group_by(continent) %>%
      select(continent, country, pop) %>%
      top_n(1, wt = desc(pop)) %>% 
      arrange(continent)
```


#### Plotting gapminder data using ggplot2
1. A scatterplot of two quantitative variables.
Look at the trend between gdpPercap and life expectancy for years after 1990
```{r}
year_data <- gapminder %>% 
      filter(year >= 1990)

p <- ggplot(year_data, aes(x=gdpPercap, y=lifeExp))
p + geom_point(aes(color=continent),alpha=0.5, size=1)
```


2. A plot of one quantitative variable
```{r}
p2 <- ggplot(gapminder, aes(lifeExp))
p2 + geom_density(aes(color=continent))
```

3. A plot of one quantitative variable and one categorical
```{r}
year2_data <- gapminder %>% 
      filter(year %in% c(1952, 2007))
p3 <- ggplot(year2_data, aes(x=factor(year), y=lifeExp))
p3 + geom_boxplot(aes(fill=factor(year))) +
      facet_grid(~ continent)
```

