---
title: "hw02_gapminder.Rmd"
author: "Fariha Khan"
date: '2017-09-24'
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




#### Work with the gapminder data we explored in class


Load Gapminder and dplyr (via tidyverse)

```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gapminder))
```

##### Explore the gapminder object:

1. Is it a data.frame, a matrix, a vector, a list?
- *The gapminder dataset is a list object and a data frame*
```{r}
str(gapminder)
```

2. What’s its class?
- *The gapminder dataset is a data frame class (made of lists)*
```{r}
class(gapminder)
typeof(gapminder)
```

3. How many variables/columns?
- *There are 6 column*
```{r}
ncol(gapminder)
```

4. How many rows/observations?
- *There are 1704 rows*
```{r}
nrow(gapminder)
```

5. Can you get these facts about “extent” or “size” in more than one way? Can you imagine different functions being useful in different contexts?
- *The str() function displays a summary of the internal structure of the object*
- *The summary() function shows the name and a summary of each variable*
- *The dim() function gives the value of rows and columns*
- *The length() function gives the number of columns and the names() fucntion outputs the names of of each column*
```{r}
str(gapminder)
summary(gapminder)
dim(gapminder)
names(gapminder)
length(gapminder)
```

6. What data type is each variable?
- *The str() function summaries the structure of the object and gives the variable type and range. The variable names and types are:*

       - country  : factor
       - continent: factor
       - year     : integer
       - lifeExp  : numerical
       - pop      : integer
       - gdpPercap: numerical
       
```{r}
str(gapminder)
```


#### Explore individual variables

##### Pick at least one categorical variable and at least one quantitative variable to explore.
I chose to explore explore the overall population by continent here.

First I used tally() to show the different columns and the number of observations in each conintent.
      
```{r tally continents}
gapminder %>% 
      group_by(continent) %>% 
      tally()
```

I used the summarize() to get the range of the minimum, maximum and average population observed in a given continent.
```{r}
gapminder %>% 
      group_by(continent) %>% 
      summarise(min_population = min(pop),
                avg_population = mean(pop),
                max_population = max(pop))
```

I tried playing with the knitr package to include a table output.
```{r, echo=FALSE, results='markup'}
library(knitr)
continent_range <- gapminder %>% 
      group_by(continent) %>% 
      summarise(min_population = min(pop),
                avg_population = mean(pop),
                max_population = max(pop))
kable(continent_range)
```


Here I added a second categorical variable and looked to find which country had the lowest population per given continent. 
```{r}
gapminder %>% 
      group_by(continent) %>%
      select(continent, country, pop) %>%
      top_n(1, wt = desc(pop)) %>% 
      arrange(continent)
```

```{r, echo=FALSE, results='markup'}
library(knitr)
low_countr <- gapminder %>% 
      group_by(continent) %>%
      select(continent, country, pop) %>%
      top_n(1, wt = desc(pop)) %>% 
      arrange(continent)

kable(continent_range)
```

### Plotting gapminder data using ggplot2

#### I used life expactancy as my main variable of interest in the following plots.

#### 1. A scatterplot of two quantitative variables.

I looked at the trend between gdpPercap and life expectancy for years **after 1990**.
I first created a filtered dataset of gapminder which only included data for years past the 90s.
The scatterplot shows an upwards trend of life expectancy with the increare is GDP per Capita. I also added distinguished the continents by colour, which shows that the Americas and Europe generally has a higher life expentancy.

```{r, echo=TRUE}
year_data <- gapminder %>% 
      filter(year >= 1990)
ggplot(year_data, aes(x=gdpPercap, y=lifeExp)) +
      geom_point(aes(color=continent),alpha=0.5, size=1) +
      labs(title= "GDP per Capita vs Life Expectancy - 1990 onwards")
```


#### 2. A plot of one quantitative variable

Here I made a density plot using geom_density to look at the life expenctancy. 
I differentiated the graph based on continent again here, which shows that different continents have their peak frequencies at various ages.
```{r, echo=TRUE}
ggplot(gapminder, aes(lifeExp)) +
      geom_density(aes(color=continent)) +
      labs(title = "Density of life expentancy - by continent")
```

#### 3. A plot of one quantitative variable and one categorical
I wanted to compare the data of the most recent and latest years here, so I created a subsetted dataframe to include data from only 1952 and 2007.
I later looked to see if there was a visable trend in the range life expectancy between these two years. I changed the year variable into a factor data type to make it discrete.
I also made the plot faceted to show the difference between the continents.
```{r, echo=TRUE}
year2_data <- gapminder %>% 
      filter(year %in% c(1952, 2007))
ggplot(year2_data, aes(x=factor(year), y=lifeExp)) +
      geom_boxplot(aes(fill=factor(year))) +
      facet_grid(~ continent) +
      scale_y_continuous(limits = c(0,100)) +
      labs(title="1952 vs 2007: Life expectancy summary stat per continent")
```

