---
title: "hw02_gapminder.Rmd"
author: "Fariha Khan"
date: '2017-09-24'
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




#### Work with the gapminder data we explored in class


Load Gapminder and dplyr (via tidyverse)

```{r cars}
library(gapminder)
library(tidyverse)
```

##### Explore the gapminder object:

1. Is it a data.frame, a matrix, a vector, a list?
```{r}
str(gapminder)
```

2. What’s its class?
```{r}
class(gapminder)
typeof(gapminder)
```

3. How many variables/columns?
```{r}
ncol(gapminder)
```

4. How many rows/observations?
```{r}
nrow(gapminder)
```

5. Can you get these facts about “extent” or “size” in more than one way? Can you imagine different functions being useful in different contexts?
```{r}
summary(gapminder)
dim(gapminder)
names(gapminder)
length(gapminder)
```

6. What data type is each variable?
```{r}
str(gapminder)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
