---
title: "Assignment 4"
author: "Fariha Khan"
date: '2017-10-09'
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load packages}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(reshape2))
```

### Activity 1
#### Data reshaping cheatsheet
reshape2 Function    | tidyr Function 
---------------------|-------------------
`melt`               |  `gather`         
`dcast`              |  `spread`         

### Activity 2
> Make a tibble with one row per year and columns for life expectancy for two or more countries.
Use knitr::kable() to make this table look pretty in your rendered homework.

I chose six countries at random for this activity, mostly because I don't know too much about them. I first created a subset of the data without applying the spread() function, so I could later plot it more easily. I used paste() to rename the column names to add the string "lifeExp" to each country column before calling kable().

```{r lifeExp 2+ countries, echo=TRUE}
actv2_lifeExp <- as.tibble(gapminder %>% 
      filter(country %in% c("Benin","Iceland","Mozambique",
                            "Sierra Leone","Venezuela", "Zambia")) %>% 
      group_by(country) %>% 
      select(country, year, lifeExp))

actv2_lifeExp_spr <- actv2_lifeExp %>% 
      spread(country, lifeExp)
names(actv2_lifeExp_spr)[2:7] <- paste(names(actv2_lifeExp_spr)[2:7], "LifeExp", sep = " ")
kable(actv2_lifeExp_spr,
      align = "c", digits = 2, padding = 1,
      caption = "Life expectancy of a given country by year")
      
```

To plot this data:
```{r plot lifeExp 2+ countries, echo=TRUE}
ggplot(data = actv2_lifeExp, aes(x = year, y = lifeExp, color = country)) +
      geom_point(aes(shape = country)) +
      geom_smooth(size = 1, method = 'lm', se = FALSE) +
      scale_color_brewer(palette = "Set2") +
      ggtitle("Linear Model of Life Expectancy from 1952 - 2007")
```

### Activity 3
> Compute some measure of life expectancy (mean? median? min? max?) for all possible combinations of continent and year. Reshape that to have one row per year and one variable for each continent. Or the other way around: one row per continent and one variable per year.


I took the variables continent, year and life expectancy here, and took measures of the minumim, mean, and average life expectancy. The first table summarizes the first 15 rows of this data.
```{r lifeExp measures, echo=TRUE, warning=FALSE}
actv3_measures <-  gapminder %>% 
      select(continent, year, lifeExp) %>% 
      group_by(year, continent) %>% 
      summarise(Min = min(lifeExp), Avg = mean(lifeExp), Max = max(lifeExp))

kable(actv3_measures[1:15,],
              align = "c", digits = 2, padding = 1,
              caption = "Measures of life expectancy per year")
```

To look at the data by one row per year and one variable for each continen, I took a spread of the minimum life expectancy.
```{r echo=TRUE}
actv3_measures %>% 
      select(year, continent, Min) %>% 
      spread(continent, Min) %>% 
      kable(align = "c", digits = 2, padding = 1,
      caption = "Minimum life expectancy per year")

```

### Activity 4
> In Window functions, we formed a tibble with 24 rows: 2 per year, giving the country with both the lowest and highest life expectancy (in Asia). 
Take that table (or a similar one for all continents) and reshape it so you have one row per year or per year * continent combination.

First input the tibble formed in class.
```{r}
actv4_tbl <- gapminder %>%
  filter(continent == "Asia") %>%
  select(year, country, lifeExp) %>%
  group_by(year) %>%
  filter(min_rank(desc(lifeExp)) < 2 | min_rank(lifeExp) < 2) %>% 
  arrange(year) %>%
  print(n = Inf)
```



