---
title: 'Homework 5: Factor and figure management'
author: "Fariha Khan"
date: '2017-10-17'
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages}
suppressPackageStartupMessages(library(forcats))
suppressPackageStartupMessages(library(singer))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(reshape2))

```
## Factor management
> Using the singer dataset

```{r}
data("singer_locations")
glimpse(singer_locations) 
```

#
### Define factor variables
> The function factor is used to encode a vector as a factor (the terms ‘category’ and ‘enumerated type’ are also used for factors)


*Taking a quick look at the dataset*
```{r}
sapply(singer_locations, function(x) length(unique(x)))
```

By looking at the dataframe, I decided that the columns for year, name and city would be good variables to cast as factors because they are discrete. Using the base as.factor() function coerces missing values to NA automatically. For forcats::as_factor(), there are limitations to the values that can be used in the function.

Note: forcats::as_factor() can't be applied to numeric objects, or NA values - producing the following errors:

 - *Error in UseMethod("as_factor") : no applicable method for 'as_factor' applied to an object of class "c('integer', 'numeric')"*

 - *Error: `idx` must contain one integer for each level of `f`*

To solve these errors I casted the year variable to characters, and explicitly marked NA values. I initially changed NA values in the catagorical variables to the string "NA" before converting to factor by using the following ifelse statement: `mutate(city_forcat_fct = ifelse(is.na(city), "NA", city), city = as_factor(city_forcat_fct))`. 
I  later found found the fct_explicit_na() function which essentially combines the two mutate statements into one.

```{r}
singer_factors <- singer_locations %>% 
      mutate(year_base_fct = as.factor(year),
             name_base_fct = as.factor(name),
             city_base_fct = as.factor(city),
             year_forcat_fct = as_factor(as.character(year)),
             name_forcat_fct = fct_explicit_na(name, na_level = "(NA)"),
             city_forcat_fct = fct_explicit_na(city, na_level = "(NA)"))

## For sanity check
glimpse(singer_factors)
singer_factors %>% 
      select(year, year_base_fct, year_forcat_fct, 
             name, name_base_fct, name_forcat_fct,
             city, city_base_fct, city_forcat_fct) %>% 
      head() %>% 
      kable(align = "c")
```



### Drop 0
Filter the singer_locations data to remove observations associated with the uncorrectly inputed  year 0. Additionally, remove unused factor levels. Provide concrete information on the data before and after removing these rows and levels; address the number of rows and the levels of the affected factors.

Looking at the singer_factor dataframe, I'm going to look at the same 3 variable is chose to manipulate in the previous step.

```{r}
with_0 <- singer_factors %>% 
      select(year_forcat_fct, name_forcat_fct, city_forcat_fct) %>% 
      summarise_all(funs(length, nlevels)) %>% 
      mutate(levels = "ALL")

dropped_0 <- singer_factors %>% 
      select(year_forcat_fct, name_forcat_fct, city_forcat_fct) %>% 
      filter(year_forcat_fct != 0) %>% 
      droplevels() %>% 
      summarise_all(funs(length, nlevels)) %>% 
      mutate(levels = "DROPPED 0")

bind_rows(with_0, dropped_0) %>% 
      rename_all(funs(gsub("_forcat_fct", "", make.names(names(with_0))))) %>% 
      column_to_rownames(var = "levels") %>% 
      kable(align = "c")
```


### Reorder levels based on knowledge from data
Reorder year in different ways

```{r}
singer_yrdrop <- singer_locations %>% 
      mutate(year = as_factor(as.character(year)),
             artist_name = fct_explicit_na(artist_name, na_level = "(NA)"),
             title = fct_explicit_na(title, na_level = "(NA)")) %>% 
      filter(year != 0) %>% 
      droplevels()

## Unordered levels
singer_yrdrop$artist_name %>%
      levels() %>% head()
      

## order by frequency
singer_yrdrop$artist_name %>% 
      fct_infreq() %>%
      levels() %>% head()

## order by reverse frequency
singer_yrdrop$artist_name %>% 
      fct_infreq() %>% fct_rev() %>% 
      levels() %>% head()
```

Reorder artist_name based on other variables using forcats::fct_reorder()
```{r}
## order artist_name by max artist_hotttnesss
fct_reorder(singer_yrdrop$artist_name, singer_yrdrop$artist_hotttnesss,
            fun = max) %>% 
      levels() %>% head()

## reverse
fct_reorder(singer_yrdrop$artist_name, singer_yrdrop$artist_hotttnesss,
            fun = max, .desc = TRUE) %>% 
      levels() %>% head()
```

Create plots to compare the difference between arrange() and fct_reorder().
I'm using the 0.75 of artist_hotttnesss as the minimum threshold for filtering artist_name

##### plotting arrange()
Plotting the values by only using arrange does not reorder the points on the graph
```{r plot arrange}
## arrange artist_name by artist_hotttnesss
top25 <- singer_yrdrop %>% 
      filter(artist_hotttnesss >= 0.75) %>% 
      arrange(desc(artist_hotttnesss)) %>% 
      ggplot(aes(x = artist_hotttnesss, y = artist_name)) +
      geom_point() +
      theme(axis.text.y = element_text(angle = 15, hjust = 1))

top25
```

##### plotting fct_reorder()
Plotting the values by only using fct_reorder() gives an graph with ordered points
```{r plot fct_reorder}
top25_reorder <- singer_yrdrop %>% 
      filter(artist_hotttnesss >= 0.75) %>% 
      ggplot(aes(x = artist_hotttnesss, y = fct_reorder(artist_name, artist_hotttnesss))) +
      geom_point() +
      theme(axis.text.y = element_text(angle = 15, hjust = 1))
top25_reorder
```

##### plotting arrange() + fct_reorder()
Plotting the values by using arrange () AND fct_reorder()  also gives an graph with ordered points
```{r plot fct_reorde + arrange}
top25_arrange_reorder <- singer_yrdrop %>% 
      filter(artist_hotttnesss >= 0.75) %>% 
      arrange(desc(artist_hotttnesss)) %>%
      ggplot(aes(x = artist_hotttnesss, y = fct_reorder(artist_name, artist_hotttnesss))) +
      geom_point() +
      theme(axis.text.x = element_text(angle = 15, hjust = 1))
top25_arrange_reorder
```

## File I/O
I'm going to save a subsetted dataframe that contains data filtered artist_hotttnesss >= 0.75,
and contained the variables for year, artist name and title.

```{r df for top25 artist_hotttnesss}
top25_hotness <- singer_yrdrop %>% 
      filter(artist_hotttnesss >= 0.75) %>% 
      select(year, artist_name, title, artist_hotttnesss) %>% 
      arrange(artist_name)
glimpse(top25_hotness)

## Save file as csv
write_csv(top25_hotness, "top25_hotness.csv")
saveRDS(top25_hotness, "top25_hotness.rds")
dput(top25_hotness, "top25_hotness.txt")
```

##### Read in top25_hotness files to see if the order of artist name remains the same

Read in using read_csv()
```{r read in csv, echo=TRUE}
csv_version <- read_csv("top25_hotness.csv")
glimpse(csv_version)
```
 - Reading the file using csv does not retain the same variable types!!
 
 - The order of the artists didn't change however
 
 
Read in using readRDS()
```{r read in rds, echo=TRUE}
rds_version <- readRDS("top25_hotness.rds")
glimpse(rds_version)
```
 - Reading the file using shows no warning messages
 
 - Looking at the glimpe of the df shows that the data types were kept the same
 
 - The order of the artists didn't change


## Visualization design

I wasn't too sure about what to do. I decided to ultimately make my own plot.
I'm going to use the full set of data, with certain variable as factors.
I'm going to use the artist names that appear most frequently and use year as changing variable to find trends.
```{r}
top_artist <- singer_yrdrop %>% 
      group_by(artist_name) %>% 
      tally() %>% 
      arrange(desc(n)) %>% 
      select(artist_name) %>% 
      head(9)
top_artists_list <- unlist(top_artist$artist_name)

plot1 <- singer_factors %>% 
      filter(artist_name %in% top_artists_list) %>% 
      ggplot(aes(x = year, y = artist_hotttnesss, color = artist_name)) +
      geom_smooth(method = lm, se = FALSE) +
      geom_jitter(size = 0.5, alpha = 0.5) +
      labs(title="Artist hotness vs year for top 9 recurring artists")+
      theme_bw()+
      scale_color_brewer(palette = "Set1") 
plot1

## This plot wasn't very informative so I decided to create another one
## 
plot2 <- singer_factors %>% 
      filter(artist_familiarity > mean(artist_familiarity),
             artist_hotttnesss > mean(artist_hotttnesss)) %>% 
      ggplot(aes(x = log10(artist_familiarity), y = log10(artist_hotttnesss))) +
      geom_smooth(method = lm, se = FALSE) +
      geom_jitter(size = 0.5, alpha = 0.2) +
      labs(title="Artist familiarity vs Artist hotttness")+
      theme_bw()+
      scale_color_brewer(palette = "Set2") 
plot2

```


## Writing figures to file
```{r}
ggsave("artistHotness_top9.png",
       plot = plot1, dpi = 500,
       width = 10, height = 7)

print(plot1)

ggsave("familliarity_vs_hotness.png",
       plot = plot2, dpi = 500,
       width = 10, height = 7)

print(plot2)
```
