---
title: "Homework 3"
author: "Fariha Khan"
date: '2017-10-03'
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#
#
##### The goal is to manipulate and explore a dataset with the dplyr package, complemented by visualizations made with ggplot2.


Load Gapminder and tidyverse (contains dplyr and ggplot)

```{r load packages}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(knitr))
```


```{r gapminder}
summary(gapminder)
```

## Task menu

### 1. Get the maximum and minimum of GDP per capita for all continents


I used the summarize() to get the range of the minimum and maximum GDP per capita observed in a given continent. The data is summarized in a table below.
```{r min gdpPercap & max gdpPercap by continent}
## Create a subset of values for min, max and average GDP per capita 
minmax_gdpPercap <- gapminder %>% 
      group_by(continent) %>% 
      summarise(min_gdpPercap = min(gdpPercap),
                avg_gdpPercap = mean(gdpPercap),
                max_gdpPercap = max(gdpPercap))
```

#### Visual representation of minimum and maximum GDP per capita by continent

```{r table and plot, results = 'asis'}
## Display table of min and max values
minmax_gdpPercap %>% 
      select(continent, min_gdpPercap, max_gdpPercap) %>% 
      kable()

## Display data in graphical format
ggplot(minmax_gdpPercap, aes(, x=continent, y=avg_gdpPercap)) +
      geom_pointrange(aes(ymin=min_gdpPercap, ymax=max_gdpPercap))

```


#
### 2. Look at the spread of GDP per capita within the continents.

#
### 3. Compute a trimmed mean of life expectancy for different years.

#
### 4. How is life expectancy changing over time on different continents?



You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
