---
title: "Homework 3"
author: "Fariha Khan"
date: '2017-10-03'
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#
#
##### The goal is to manipulate and explore a dataset with the dplyr package, complemented by visualizations made with ggplot2.


Load Gapminder and tidyverse (contains dplyr and ggplot)

```{r load packages}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(grid))
suppressPackageStartupMessages(library(gridExtra))


```


```{r gapminder}
summary(gapminder)
```

## Task menu

### 1. Get the maximum and minimum of GDP per capita for all continents


I used the summarize() to get the full range GDP per capita observed in a given continent, and created a subset so I could use it for later questions. 



```{r}
summary_gdpPercap <- gapminder %>% 
      group_by(continent) %>% 
      summarise(Minimum = min(gdpPercap),
                Q1 = quantile(gdpPercap, probs=0.25),
                Median = median(gdpPercap),
                Q3 = quantile(gdpPercap, probs=0.75),
                Maximum = max(gdpPercap),
                Average = mean(gdpPercap))
```

Here, the minimum and maximum GDP per capita is summarized in a table below by continent.
```{r min gdpPercap & max gdpPercap by continent, echo=TRUE, results='markup'}
## Select values for max and min 
## Display values in a table
summary_gdpPercap %>% 
      select(continent, Maximum, Minimum) %>% 
      kable()
```

```{r, echo=TRUE}

## Display values in a table
ggplot(summary_gdpPercap, aes(x=continent, color = continent)) +
       geom_errorbar(aes(ymin=Minimum, ymax=Maximum))+
       ggtitle("Max and Min GDP per Capita") +
       ylab("GDP per Capita")
```

> I've never used geom_linerange before - it seemed to be the best way to graphically show only two values per variable.


#
### 2. Look at the spread of GDP per capita within the continents.
```{r GDP per capita spread, echo=TRUE, results='markup'}
kable(summary_gdpPercap)
```

I chose to use a histogram to show the distribution of GDP per capita. I used the original gapminder dataset to show the full distribution of the values.

To explore whether there is any trend correlating to year and GDP per capita, I used a a combination of a point graph overlapped with a linear fit graph.
```{r plot, fig.height=5, fig.width=10}
gdppc_hist <- ggplot(gapminder, aes(x=gdpPercap, fill = continent)) + 
      geom_histogram(binwidth=500) +
      facet_wrap(~ continent, scales ="free", nrow = 1) +
      scale_fill_brewer(palette = "Dark2") +
      theme(legend.position="bottom")

gdppc_line <- ggplot(gapminder, aes(year, gdpPercap, colour=continent)) +
      geom_point(alpha=0.5, size=0.1) +
      geom_smooth(method="lm") +
      scale_color_brewer(palette = "Dark2") +
      facet_grid(~continent) +
      theme(legend.position="top")


grid.arrange(gdppc_line,gdppc_hist, ncol=1,
             top = textGrob("GDP per capita by continent"))
```



```{r fig.height=5, fig.width=10,}

```


#
### 3. Compute a trimmed mean of life expectancy for different years.
*Couldn't think of a cool analysis*


#
### 4. How is life expectancy changing over time on different continents?
I did a simple analysis of looking at the change in life expectancy (five number summary), based on year and continent. 
```{r, results='markup'}
summary_lifeExp <- gapminder %>% 
      group_by(continent, year) %>% 
      summarise(Minimum = min(lifeExp),
                Q1 = quantile(lifeExp, probs=0.25),
                Median = median(lifeExp),
                Q3 = quantile(lifeExp, probs=0.75),
                Maximum = max(lifeExp),
                Average = mean(lifeExp)) 

kable(head(summary_lifeExp))
```

I chose to plot the data by using a line-type graph overlaying geom_points to show the positive correlation between year and life expectancy. This trend seems to be consisent through each continent.
```{r}
ggplot(gapminder, aes(x=year, y=lifeExp, shape = continent, color=continent)) +
      geom_point(aes(group=continent),
                 size = 1, position = position_dodge(width = 2), alpha =0.6) +
      geom_smooth(aes(group=continent)) +
      scale_color_brewer(palette = "Set1") +
      theme_bw()
```

You can also embed plots, for example:

```{r pressure, echo=FALSE}
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
